---
title: "hw09-lsloboda"
output:
   github_document:
      toc: true
always_allow_html: yes
---

# Automating a Pipeline

## Objective

The objective of the assignment is to add an R file and an Rmd file to a pre-seeded pipeline, originally found [here]("https://github.com/STAT545-UBC/make-activity").

## Method

I added the following files to the pipeline:

* letterCount.R, which counts the number of each letter in word.txt
* word_analysis_report.Rmd, which outputs an .html and .md report of the results of letterCount.R

I updated the Makefile by:

* Adding instructions to clean the new files
* Adding instructions to run the new files
* Adding new dependencies, where necessary

## Scripts

The script letterCount.R counts the number of letters in all words in word.txt and write them to a .tsv as a data frame:
```{r}
#words <- readLines("words.txt")
#letterCount <- sapply(letters, function(x) x<-sum(x == unlist( strsplit( words, "" )) ))
#str(letterCount)

#write.table(letterCount, "letterCount.tsv", sep = "\t") #automatically converts to a data frame
```

The markdown file word_analysis_report.Rmd output the results summary from letterCount.R:
```{r}
#title: "word_analysis_report"
#output:
#  html_document:
#    keep_md: yes

#suppressPackageStartupMessages(library(readr))
#suppressPackageStartupMessages(library(knitr))
#suppressPackageStartupMessages(library(tidyverse))

#letterCount <- read.table("letterCount.tsv")
#kable(letterCount, col.names = "Count")
```

## Makefile

The makefile is updated with instructions regarding the new files, including adding new dependencies, as well as running and cleaning the new files:
```{r}
#all: report.html word_analysis_report.html

# clean:
# 	rm -f words.txt histogram.tsv histogram.png report.md report.html letterCount.tsv word_analysis_report.html word_analysis_report.md
# 
# report.html: report.rmd histogram.tsv histogram.png
# 	Rscript -e 'rmarkdown::render("$<")'
# 
# histogram.png: histogram.tsv
# 	Rscript -e 'library(ggplot2); qplot(Length, Freq, data=read.delim("$<")); ggsave("$@")'
# 	rm Rplots.pdf
# 
# histogram.tsv: histogram.r words.txt
# 	Rscript $<
# 
# #words.txt: /usr/share/dict/words
# #	cp $< $@
# 
# words.txt:
# 	Rscript -e 'download.file("http://svnweb.freebsd.org/base/head/share/dict/web2?view=co", destfile = "words.txt", quiet = TRUE)'
# 
# letterCount.tsv: letterCount.R words.txt
# 	Rscript $<
# 
# word_analysis_report.html: word_analysis_report.Rmd letterCount.tsv
# 	Rscript -e 'rmarkdown::render("$<")'
```

Although very simple tasks are used in this example, automating a pipeline is very handy to save time when it is necessary to re-run the same scripts often and in a particular order.

## Resources

I had trouble getting make to work on Windows. It ended up being a PATH problem, which I fixed by uninstalling and re-installing RTools with the automatic PATH setup, thanks to help from this forum post:
https://support.rstudio.com/hc/en-us/community/posts/209092218-Make-is-not-recognized-as-an-internal-or-external-command

https://support.rstudio.com/hc/en-us/articles/200486498-Package-Development-Prerequisites

https://cran.rstudio.com/bin/windows/Rtools/

http://www.sthda.com/english/wiki/fast-writing-of-data-from-r-to-txt-csv-files-readr-package


